Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-04 files/Overview.tex
overview v2

\subsubsection{Chapter 1}
This dissertation is a literature study to investigate the validity of different linear models for beams and plates. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The simplest model is the Timoshenko model. The other models are a two-dimensional elastic beam model, a three-dimensional elastic beam model and a Reissner-Mindlin plate model. The aim in this dissertation to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exists, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see Chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the modal solutions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In Chapter 1, the models of the dissertation is given (they are given in a dimensionless form).  Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation. Also the variational forms are derived which enables us to see the similarity between the models.

\subsubsection{Chapter 2}
First, a variational form for a model problem of a cantilever Timoshenko beam given. This model problem in variational form is then extended to complete function spaces, and in turn a weak variational problem is obtained. This weak variational problem is used as an example to explain the main theory of the chapter, from the article \cite{VV02}. In this article, the authors present a general weak variational form. The weak variational form of all the model problems of this dissertation are special cases of this general weak variational problem. Therefore the assumptions and results can be formulated and applied to all the applications. The article gives four assumptions. Under these assumptions it is shown that the general vibration problem has a unique solution using semi-group theory. Next the theory is applied to the example weak variational problem for the cantilever Timoshenko beam, by showing that the assumptions hold.

Finally the concept of modal analysis is introduced, which is fundamental to this dissertation. The idea of modal analysis is explained by hands of another example, again using a cantilever Timoshenko beam. Then the general case is discussed, following \cite{CVV18}. Given a general vibration problem, an eigenvalue problem and a differential equation is derived. An additional assumption is introduced by \cite{CVV18}. Under this assumption, the eigenvalue problem has infinite many solutions, and that the solution of the general vibration problem is an infinite series of the eigenvectors and general Fourier coefficients. Solving the differential equation, shows that these Fourier coeffiencients are periodic and a formal series solution for the general vibration problem (which is a boundary value problem) is obtained. This formal series solution is then shown to also be valid for a general initial value problem. This general case is applicable to all the models in this dissertation.

Being able to express the solution of a model problem as a linear combination of the eigenvalues and eigenfunctions in a formal series solution, is fundamental to this dissertation. This means that the solutions of models will compare well if the eigenvalues and eigenfunctions compare well. The validity of the formal series solution is also a requirement as it ensures that the comparisons remain valid, as long as the models are disturbed in the same way.

\subsubsection{Chapter 3}
In this chapter, some theory for the Finite Element Method is discussed. This chapter contains two parts, that covers two different results. The first result is the convergence of the Galerkin approximation of a general vibration problem. This result is presented in the article \cite{BV13}. The second result concerns the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. This result is presented in the textbook \cite{SF97}. In the first section, the authors continue with the general vibration problem that is used in Chapter 2. The general vibration problem has a solution as shown in Chapter 2. The Galerkin Approximation is derived from the general vibration probelm and is rewritten into a system of ordinary differential using the Finite Element Method. This oridinary differential equation can be proven to have a unique solution. The main results of \cite{BV13} shows that this solution of the Galerkin Approximation converges to the solution of the general vibration problem. The approach of the authors is to calculate the error estimates. The error is split into two parts, a semi-discrete and a fully discrete problem. Two assumptions are added to the assumptions from Chapter 2. Using these assumptions, the authors obtain error estimates for the semi-discrete and fully discrete problems.

@@

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this dissertation, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The examples shows the use of the method, and is preperation for chapter 6 where these results are required for the comparison of the models. The next section is a short discussion on a suspended beam model, afterwards an important article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors suspended the beam, and vribrated the beam to induce a free-free beam and measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and look at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.

\subsubsection{Chapter 5}
In this chapter, the Finite Element Method is applied to cantilever two-dimensional,and three-dimensional cantilever beams and a cantilever Reissner-Mindlin plate. The aim of this section is to obtain an algorithm to calculate the eigenvalues and eigenfunctions of the models. The results of these formulas will be used in chapter 6. The Finite Element Method is not applied to the Timoshenko beam theory, as chapter 4 provides an alternative method. For all the models, the Finite Element Method is applied using bi-cubic or tri-cubic basis functions to improve the rate of convergence and reduce the processing required to obtain accurate results. Each section ends with a eigenvalue problem for the models that can be easily applied to a computer program to calculate the eigenvalues and eigenvectors for the models. These algorithms will be used to obtain the results in chapter 6.

\subsubsection{Chapter 6}
In this chapter, the main comparisons of the models are conducted. The first section is a discussion of the article \cite{LVV09}. Each section in this chapter follows the same structure as this article. In this article, the authors investigate the validity of a cantilever Timoshenko beam models, by comparing it to a two-dimensional cantilever Timoshenko beam model. To make this comparison, the authors compare the eigenvalues and eigenfunctions of the models. But since the two-dimensional model is more complex, there are `non-beam type' eigenvalues. These are eigenvalue that are not in the Timoshenko beam theory, and does not relate to beam type problems. To be able to find the beam type eigenvalues that can be compared, the authors compare the mode shapes. This allows the eigenvalues to be matched up. The results of the article show that the Timoshenko beam model compares very well to the two-dimensional beam. The comparison improves when the application is for a long and slender beam. The results of the article are verified and more results are added to extend the article and also to be using the the following sections. But in real-world applications, a beam is a three-dimensional model. Therefore it would be better to use a three-dimensional model to investigate the validity of the Timoshenko beam theory. The authors of \cite{LVV09} mentions this, and suggests using the two-dimensional model as an intermediate step to avoid complexities. So in the second section of this chapter, the validity of the two-dimensional model is investigated using the three-dimensional model as a reference. The results show that the two-dimensional model also compares very well to the three-dimensional model. Similar to the one-dimensional case, the results improve for a long and slender beam, and also if the width of the three-dimensional beam is not larger than the height of the beam. When increasing the width past the size of the height, the two-dimensional model is less accurate. This shape does however bring into question the use of a beam model, and something like a plate model might be better suited. In the last section, the validiy of the Reissner-Mindlin plate model is investigated. Similar to how the Timoshenko beam theory improves on the Euler-Bernoulli beam theory, the Reissner-Mindlin plate theory improves on the classical plate theory. The numerical results do show that this Reissner-Mindlin plate model compares well to a three-dimensional plate model, also when the dimensions of the models are realistic for application.

>>>>

\subsubsection{Chapter 1}
This dissertation is a literature study to investigate the validity of different linear models for beams and plates. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The simplest model is the Timoshenko model. The other models are a two-dimensional elastic beam model, three-dimensional elastic beam and plate models and a Reissner-Mindlin plate model. The aim in this dissertation to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exists, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see Chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the modal solutions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In Chapter 1, the models of the dissertation is given (they are given in a dimensionless form).  Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation. The variational forms are also derived which enables us to see the similarity between the models and the general theory.

\subsubsection{Chapter 2}
First, a variational form for a model problem of a cantilever Timoshenko beam given. This model problem in variational form is then extended to complete function spaces, and a weak variational problem is obtained. This weak variational problem is used as an example to explain the main theory of this chapter which is from the article \cite{VV02}. In this article, the authors present a general weak variational form. The weak variational form of all the model problems of this dissertation are special cases of this general weak variational problem. Therefore the assumptions and results can be formulated and applied to all the applications in this dissertation. The article gives four assumptions. Under these assumptions it is shown that the general vibration problem has a unique solution using semi-group theory. The theory is then applied to the example by proving that the assumptions hold.

Finally the concept of modal analysis is introduced, which is fundamental to this dissertation. First idea of modal analysis is explained by hands of another example, again using a cantilever Timoshenko beam. Then the general case is discussed that follows the article \cite{CVV18}. Given a general vibration problem, it is split into two problems by introducing a trial solution. It can then be verified that these two problems are the eigenvalue problem and a differential equation. An additional assumption (additional to the assumptions of \cite{VV02}) is introduced by \cite{CVV18}. Under this assumption, the eigenvalue problem has infinite many solutions, and the solution of the general vibration problem is an infinite series of the eigenvectors and general Fourier coefficients. Solving the differential equation, shows that these Fourier coefficients are periodic and a formal series solution for the general vibration problem (which is a boundary value problem) is obtained. This formal series solution is then shown to also be valid for a initial value problem. This general case is also applicable to all the models in this dissertation.

This theory is crucial for this dissertation. It ensures that the solutions of the models will compare well if the eigenvalues and eigenfunctions compare well. Therefore for the comparisons in Chapter 4 and 6, it is only required to compare the eigenvalues and eigenfunctions. The validity of the formal series solution is also important as it ensures that the comparisons remain valid, as long as the models are disturbed in the same way.

\subsubsection{Chapter 3}
\textcolor{red}{Nie seker hoekom prof gese het nie BV13 nie? Ook nie seker hoekom SF97 dood getrek was nie.}

In this chapter, some theory for the Finite Element Method is discussed. This chapter contains two parts, that covers two different theoretical results. The first is on the convergence of the Galerkin approximation of a general vibration problem. This theory is presented in the article \cite{BV13}. The second part concerns the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. This theory is presented in the textbook \cite{SF97}.

In the first section, the authors continue with the general vibration problem that is used in Chapter 2. The general vibration problem has a solution as shown in Chapter 2. The Galerkin Approximation is derived from the general vibration probelm and is rewritten into a system of ordinary differential using the Finite Element Method. This oridinary differential equation can be proven to have a unique solution. The main results of \cite{BV13} shows that this solution of the Galerkin Approximation converges to the solution of the general vibration problem. The approach of the authors is to calculate the error estimates. The error is split into two parts, a semi-discrete and a fully discrete problem. Two assumptions are added to the assumptions from Chapter 2. Using these assumptions, the authors obtain error estimates for the semi-discrete and fully discrete problems.

@@

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this dissertation, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The examples shows the use of the method, and is preperation for chapter 6 where these results are required for the comparison of the models.

The next section is a short discussion on a suspended beam model, afterwards an important article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors suspended the beam, and vribrated the beam to induce a free-free beam and measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and investigates at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.

\subsubsection{Chapter 5}
In this chapter, the Finite Element Method is applied to cantilever two-dimensional elastic body, cantilever three-dimensional elastic body and a cantilever Reissner-Mindlin plate. The aim of this section is to obtain an algorithm to calculate the eigenvalues and eigenfunctions of the models. The Finite Element Method is not applied to the Timoshenko beam theory, as chapter 4 provides an alternative method. For all the models, the Finite Element Method is applied using bi-cubic or tri-cubic basis functions to improve the rate of convergence and reduce the processing required to obtain accurate results. Each section ends with a eigenvalue problem for the models that can be easily applied to a computer program to calculate the eigenvalues and eigenvectors for the models. This is in preparation of Chapter 6 where the eigenvalues and eigenfunctions are calculated and compared.

\subsubsection{Chapter 6}
In this chapter, the main comparisons of the models are conducted. All of the sections follow the same idea and struture as the article \cite{LVV09}.

The first section is a discussion of the article \cite{LVV09}. In this article, the authors investigate the validity of a cantilever Timoshenko beam model, by comparing it to a two-dimensional cantilever Timoshenko beam model. The authors compare the models by comparing the eigenvalues and eigenfunctions (see Chapter 2 on modal analysis). The two-dimensional model is more complex and there are eigenvalues that are not shared between the models. The authors use the mode shapes to match up the eigenvalues. The eigenvalues matching the eigenvalues of the Timoshenko beam model are referred to by the authors as beam-type eigenvalues. The authors also consider different shapes of beams, from a short fat beam to a long slender beam. The results show that the models compare very well, even for a short and fat beam. As the beam gets more slender, the comparison improves. In this section some of the results of the article are discussed and replicated.

But in real-world applications, a beam is a three-dimensional model. Therefore it would be more suited to use a three-dimensional model to investigate the validity of the Timoshenko beam theory. This is mentioned by the authors of \cite{LVV09}. Their suggestion is to use the two-dimensional model as an intermediate step, to avoid complexities. Therefore the validity of the two-dimensional model is investigated, using a three-dimensional beam model as a reference. Again the results so that the comparison relies on the shape of the beams. The two-dimensional model compares well to the three-dimensional beam if the beam is not very wide.

If the width of the beam is very large, the use of a beam model can be questioned. A plate model might be more suited. Therefore the last section of this chapter investigates the validity of the Reissner-Mindlin plate model. The Reissner-Mindlin plate model is compared to a three-dimensional plate model. The results show that the Reissner-Mindlin plate model compares well to the three-dimensional plate model. This is also true for a wide plate.

Overall the models compare very well, as long as the shape of the model is realistic.
